---
title: "Homework"
format: html
editor: visual
---

## Project Description

This project is one of the building blocks for a bigger project of my PhD work that looks into the effects of spring timing on the breeding success of alpine birds in central Norway. The study has data from 2019 to 2025 where we have field survey data showing some years of good breeding success and some years of failure.

In order to effectively quantify the timing of spring first, I use camera data from the field sites (the cameras are set up across 7 different plots within the study site). We use

Year as a fixed effect because the purpose is to see the difference each year. This is at the central of my question later part of my project looks into how t

## Analysis

Step 1: Data Manipulation

```{r}
#| echo: false
#| ## load necessary libraries##
library(readxl)
library(stringr)
library(ggplot2)
library(brms)
library(cmdstanr)
library(tidyverse)
library(tidybayes)

## Create a function to make sure data is in right format 
clean_daily_file <- function(file_path) {
  df <- readxl::read_excel(file_path)
  file_name <- basename(file_path)
  year <- stringr::str_extract(file_name, "20\\d{2}") ## extract from file name
  df_clean <- df %>%
    mutate(
      Plot = as.factor(stringr::str_extract(MemCard, "\\d{4}$")), # extract from sd card name
      Date = as.Date(Date),
      meanTempC = as.numeric(meanTempC),
      JulianDay = as.integer(format(Date, "%j")), # get from date
      Year = as.factor(year),
      SourceFile = file_name,
      meanGreenFraction = meanGreenFraction / 100, ## convert original percent value to proportion out of 100
      N_pixel = 10000, ## randomly estimated 
      n_green = as.integer(round(meanGreenFraction * N_pixel)) ## convert to successful number of trials as integer 
    )
  return(df_clean)}
files <- list.files(
  path = "/cloud/project/Greeness",
  pattern = "^dailyGreen_20\\d{2}_RypeLierne.*\\.xlsx$",
  full.names = TRUE)

combined_data <- bind_rows(lapply(files, clean_daily_file))
head(combined_data)
```

You can add options to executable code like this

```{r}

Model_3 <- brms::brm(n_green | trials(N_pixel) ~ Year + s(JulianDay, by = Year) + (1|Plot), 
                     data = combined_data, 
                     family = beta_binomial (link = "logit"), 
                     chains = 3,
                     backend = "cmdstanr", 
                     iter = 2000, warmup = 500, 
                     seed = 123)
saveRDS(Model_3, "greening_M3.RDS") ## so u dont need to rerun again
summary(Model_3) 
#bulk and Tail TSS is sample effect size and must be higher than 1000 and Rhat should be 1 if converged well 
fixef(Model_3) ## detailed value of estimates 
coef(Model_3) ## group level effect
plot(Model_3) ## should be a fuzzy caterpillar pattern for good convergence 
pp_check(Model_3) ## check if model predicts your data accurately 
```

The `echo: false` option disables the printing of code (only output is displayed).
